Locally-Oriented Programming:\\ A Simple Programming Model for Stencil-Based Computations on Multi-Level Distributed Memory Architectures

### 2015-01-26 papers/LOPe/coarray-extensions.tex
Working on subimage description.

Consider a simple average over a five point stencil in two dimensions.  The algorithm updates the
stencil interior point $A(0,0)$ as shown in the following:

The

\begin{verbatim}
   real, allocatable, HALO(1:*:1,1:*:1) :: U(:,:)[:]

   !! Allocate normal coarray memory
   !  -------------------------------------------
   allocate(U(M,N)[*])
\end{verbatim}

\begin{verbatim}
   device = get_subimage(1)          ! THIS_IMAGE() if device unavailable

>>>>

Consider the \texttt{Laplace} concurrent function whose implementation was provided above.
In this section we demonstrate how this function can be called in a context that allows
full access to all of the enterior elements of the array, as well as array element within
the halo boundary region.  Topics highlighted in this section are: 1. Distributed memory
array allocation; 2. The placement or location(s) (possibly multiple) of the memory
allocated for the distributed array; 3. Remote memory transfer; and 4. Remote execution.
This description is within the context of a Domain Specific Language (DSL) extending
Fortran 2008.  These specific extensions are called CAFe, (Coarray Fortran extensions).

We start with the declaration of an array (to be allocated later),
\begin{verbatim}
  real, allocatable, HALO(1:*:1,1:*:1), dimension(:,:), codimensions(:,:) :: U
\end{verbatim}
with an explicit halo size and with two local dimensions (rank) and two distributed memory
codimensions (corank).  The corank of the array is chosen to be identical to the rank of
the array so that the logical process topology aligns is a way that allows a natural
halo exchange between logically neighboring processes.  For example, if my process
location is \texttt{[pcol,prow]}, then the right-hand halo for my local array \texttt{U}
can be obtained by the assignment \texttt{U(M+1,:) = U(1,:)[pcol-1,prow]} where the size
and cosize of \texttt{U} are given by the allocation statement,
\begin{verbatim}
  allocate(U(0:M+1,0:N+1)[MP,*])
\end{verbatim}

This allocation statement specifies that (given the 1 element halo size provided earlier for
\texttt{U}) the local, left halo is \texttt{U(0,:)}, the right is \texttt{U(M+1,:)},
the bottom is \texttt{U(:,0)} and the top is \texttt{U(:,N+1)}, leaving the enterior
region to be \texttt{U(1:M,1:N)}.  In this allocation statement the total number of
process rows \texttt{NP} can be obtained at runtime and is not a compile time parameter
because the actual number of participating processes (in Fortran called images) is
variable, depending on how many processes are requested at startup.  In this discussion,
it is \emph{assumed} that there are no holes in the logical processor topology, thus
$MP*NP = P$, where $P$ is the total number of participating processes (images).

So far --- except for the \texttt{HALO} attribute --- this is totally standard Fortran (2008).  For CAFe,
we extend the concept of Fortran images to be hierachical.  By this we mean that each image \emph{may}
have a subimage (or subimages), but this subimage is not visible to other normal Fortran images.
Subimages also execute differently than normal images.  Subimages are task based while images all
execute a single program but with different (multiple) data (SPMD).  A task can be given to a subimage,
but execution on the subimage terminates once the task is finished.  Memory on a subimage, however, is
permanent and must be specifically allocated and deallocated.

One obtains a subimage by executing,
\begin{verbatim}
   device = get_subimage(1)          ! THIS_IMAGE() if device unavailable
\end{verbatim}

\begin{verbatim}
